Apparent preferences for cognitive effort fade when multiple forms of effort

### 2021-06-09 Manuscript.Rmd
fixed fig width and added data / code availability sections

## figures at the end of the manuscript?
figsEnd <- F

@@

@@

```{r New General prop & earnings plots (Fig. 2), eval = figsEnd, fig.width = 10, fig.height = 4, fig.align = "center", fig.cap = "Overall choice behavior in Experiment 1. A: Proportion accepted per cost. B: Total number of dollars earned by each group by the end of the experiment (not including the participation reward). C: The relationship between proportion accepted and total earned. Consistent with the foraging design, participants who over and under accepted earned the least. These results suggest that people were willing to pursue cognitively effortful actions at the expense of earnings."}
# plots generated in 16.2.1. and 16.1.3.
# propPlot <- (prop1 | earnings1 | earnFitplot_btw) + 
#     plot_annotation(tag_levels = 'A') &
#     theme(plot.tag = element_text(size = 20, face = "bold"))
# ggsave("./Images/FIG2_props.png", propPlot, device = "png", height = 4, width = 10)
include_graphics("./Images/FIG2_props.png")

@@

```{r New mixed effects plots (Fig. 3), eval = figsEnd, out.width = "70%", fig.align = "center", fig.cap = "Comparison of acceptance rates across cost types for each combination of reward amount and handling time. A: Mean acceptance rate in each group for every combination of handling time and reward (with SEM). Grey points indicate the reward-maximizing acceptance rate for each combination, which was always either 0 or 1. B: Matrix of coefficients that resulted from switching the reference cost condition in a mixed-effects logistic regression model. Each element represents how much more likely a group was to accept offers compared to the reference condition (row). White asterisks denote the p-value for significant comparisons (one, two, and three askterisks correspond to p-values under 0.05, 0.01, and 0.001, respectively). C: AIC and BIC values for a mixed-effects model comparison, ranging from a baseline intercept-only model to a full-interaction model. Both metrics yielded similar values across models."}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG3_btwlogistic.png")
```

@@

```{r New Survival and prepost plots (Fig. 4), eval = figsEnd, fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "Choice consistency and stability in Experiment 1. A: Survival curve plots of trial rejection times. B: Distributions of trial rejection times within the first 1 s of the handling time. C: Acceptance rate in the second half of the session as a function of acceptance rate in the first half of the session. Dashed lines indicate the mean for each group. D: proportion accepted over time per group."}
# individual plots generated in 16.3.1. and 16.3.2.
# to recreate the combined plot and save it:
# historyPlot_btw <- (suppressWarnings(survPlot_btw) | RTplot_btw) / (prepost_btw$scatterplot | prepost_btw$propBlocks) + 
#   plot_annotation(tag_levels = 'A') & 
#   theme(plot.tag = element_text(size = 20, face = "bold"))
# ggsave("./Images/FIG4_survivalbtw.png", historyPlot_btw, device = "png", height = 8, width = 10)
include_graphics("./Images/FIG4_survivalbtw.png")
```

@@

```{r New Prepost mixed effects plot (Fig. 5), eval = figsEnd, fig.align = "center", fig.cap = "Acceptance behavior over time in Experiment 1. A: Offer acceptance rates per handling time, reward amount, cost type, and session half, showing that effortful demands (Cognitive and Physical) were less frequently accepted over time, especially for the longest handling time (14 s). B: Mixed-effects logistic regression model coefficients denoting the comparison of acceptance rates across cost conditions for each half of the experimental session separately (mirrored along the diagonal). Even though acceptance rates decreased in the effort groups, the differences in acceptance rates across cost conditions were preserved."}
# plots in 16.3.2
# ggsave("./Images/prepost_btw.png", hrprepostPlot_btw, device = "png", height = 6, width = 6)
include_graphics("./Images/FIG5_prepostbtw.png")
```

@@

```{r New Earnings/prop completed/ME plots (Fig. 6), eval = figsEnd, fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "Within-subject acceptance behavior. A: Quadratic relationship between trial completions and total earnings. B: Proportion of acceptances of each reward per cost condition. C: Matrix of mixed-effects coefficients showing comparisons among cost conditions (while controlling for reward amount). Cognitive and Physical trials were compared to one another and were each compared to Wait trials from the same block (Wait-C and Wait-P, respectively). We omitted comparisons to Wait trials in the opposite block type (gray cells), which have little interpretability and were not addressed in our predictions. Asterisks mark statistically significant comparisons. D: AIC and BIC values for models of increasing complexity (AllMain refers to the a priori model which included main effects of cost condition and reward offer amount.)"}
# earnFitplot_wth, hrPlot_wth, mixLogis_compare_wth$plot
#ggsave("./Images/aic_wth.png", mixLogis_compare_wth$plot, device = "png", height = 5, width = 5)
include_graphics("./Images/FIG6_wthlogistic.png")
```

@@

```{r New Survival and prepost plots (Fig. 7), eval = figsEnd, fig.width = 8, fig.height = 7, fig.align = "center", fig.cap = "Choice consistency and stability in Experiment 2. A: Survival curves showing trial rejection times for each cost condition. B: Distribution of trial rejection times within the offer window. C: Acceptance rate in the last two blocks as a function of acceptance rate in the first two blocks. Dashed lines indicate the mean for each cost. D: Acceptance rates in the three successive occurrences of each block type."}
# plots generated in 16.3.1. and 16.3.3.
# historyPlot_wth <- (suppressWarnings(survPlot_wth) | suppressWarnings(RTplot_wth)) / (prepost_wth$scatterplot | prepost_wth$propBlocks) +
#     plot_annotation(tag_levels = 'A') &
#     theme(plot.tag = element_text(size = 20, face = "bold"))
#ggsave("./Images/FIG7_survivalwth.png", historyPlot_wth, device = "png", height = 8, width = 10)
include_graphics("./Images/FIG7_survivalwth.png")
```

@@

```{r New BIC and parameter comparison (Fig. 9), eval = figsEnd, fig.align = "center", fig.height = 5, fig.width = 7, fig.cap = "Computational modeling results. A: iBIC values for four candidate models of choice behavior for both experiments. Lower values indicate better fit. B: Estimates of the bias parameter c-demand for participants in the Physical, Cognitive, and Wait cost conditions of Experiment 1. Negative values suggest the Cognitive group acted as if cognitive demands attenuated the perceived opportunity cost of the delay. C: Estimates of other model parameters indicated consistent foraging behavior across conditions and experiments. D: Synthetic behavior for Experiment 2 generated by the winning model. Error bars correspond to the empirical mean acceptance rate ± SEM per cost type, whereas lines and shaded areas show the model-generated mean acceptance rate and its SEM, respectively. The model's behavior recapitulated the initial apparent preference for cognitive effort and the eventual convergence of acceptance rates by the end of the session."}
# modelPlot <- (bicPlot + costPlot) / allFitsPlot
# ggsave("./Images/FIG9_modelresult_part.png", modelPlot, device = "png", height = 8, width = 11)
include_graphics("./Images/FIG9_modelresults.png")
```

>>>>

## figures at the end of the manuscript?
figsEnd <- T

@@

## Data availability

The data obtained from human participants are available at https://github.com/ctoroserey/Cost_studies

## Code availability

The code used to perform all analyses can be found in https://github.com/ctoroserey/Cost_studies

@@

```{r New General prop & earnings plots (Fig. 2), eval = figsEnd, out.width = "100%", fig.align = "center", fig.cap = "Overall choice behavior in Experiment 1. A: Proportion accepted per cost. B: Total number of dollars earned by each group by the end of the experiment (not including the participation reward). C: The relationship between proportion accepted and total earned. Consistent with the foraging design, participants who over and under accepted earned the least. These results suggest that people were willing to pursue cognitively effortful actions at the expense of earnings."}
# plots generated in 16.2.1. and 16.1.3.
# propPlot <- (prop1 | earnings1 | earnFitplot_btw) + 
#     plot_annotation(tag_levels = 'A') &
#     theme(plot.tag = element_text(size = 20, face = "bold"))
# ggsave("./Images/FIG2_props.png", propPlot, device = "png", height = 4, width = 10)
include_graphics("./Images/FIG2_props.png")

@@

```{r New mixed effects plots (Fig. 3), eval = figsEnd, out.width = "80%", fig.align = "center", fig.cap = "Comparison of acceptance rates across cost types for each combination of reward amount and handling time. A: Mean acceptance rate in each group for every combination of handling time and reward (with SEM). Grey points indicate the reward-maximizing acceptance rate for each combination, which was always either 0 or 1. B: Matrix of coefficients that resulted from switching the reference cost condition in a mixed-effects logistic regression model. Each element represents how much more likely a group was to accept offers compared to the reference condition (row). White asterisks denote the p-value for significant comparisons (one, two, and three askterisks correspond to p-values under 0.05, 0.01, and 0.001, respectively). C: AIC and BIC values for a mixed-effects model comparison, ranging from a baseline intercept-only model to a full-interaction model. Both metrics yielded similar values across models."}
#hrPlot_btw / (ABIC_plot_btw | ME_matrix_btw) # doesn't work with corrplot.. oh well
include_graphics("./Images/FIG3_btwlogistic.png")
```

@@

```{r New Survival and prepost plots (Fig. 4), eval = figsEnd, out.width = "95%", fig.align = "center", fig.cap = "Choice consistency and stability in Experiment 1. A: Survival curve plots of trial rejection times. B: Distributions of trial rejection times within the first 1 s of the handling time. C: Acceptance rate in the second half of the session as a function of acceptance rate in the first half of the session. Dashed lines indicate the mean for each group. D: proportion accepted over time per group."}
# individual plots generated in 16.3.1. and 16.3.2.
# to recreate the combined plot and save it:
# historyPlot_btw <- (suppressWarnings(survPlot_btw) | RTplot_btw) / (prepost_btw$scatterplot | prepost_btw$propBlocks) + 
#   plot_annotation(tag_levels = 'A') & 
#   theme(plot.tag = element_text(size = 20, face = "bold"))
# ggsave("./Images/FIG4_survivalbtw.png", historyPlot_btw, device = "png", height = 8, width = 10)
include_graphics("./Images/FIG4_survivalbtw.png")
```

@@

```{r New Prepost mixed effects plot (Fig. 5), eval = figsEnd, out.width = "95%", fig.align = "center", fig.cap = "Acceptance behavior over time in Experiment 1. A: Offer acceptance rates per handling time, reward amount, cost type, and session half, showing that effortful demands (Cognitive and Physical) were less frequently accepted over time, especially for the longest handling time (14 s). B: Mixed-effects logistic regression model coefficients denoting the comparison of acceptance rates across cost conditions for each half of the experimental session separately (mirrored along the diagonal). Even though acceptance rates decreased in the effort groups, the differences in acceptance rates across cost conditions were preserved."}
# plots in 16.3.2
# ggsave("./Images/prepost_btw.png", hrprepostPlot_btw, device = "png", height = 6, width = 6)
include_graphics("./Images/FIG5_prepostbtw.png")
```

@@

```{r New Earnings/prop completed/ME plots (Fig. 6), eval = figsEnd, out.width = "95%", fig.align = "center", fig.cap = "Within-subject acceptance behavior. A: Quadratic relationship between trial completions and total earnings. B: Proportion of acceptances of each reward per cost condition. C: Matrix of mixed-effects coefficients showing comparisons among cost conditions (while controlling for reward amount). Cognitive and Physical trials were compared to one another and were each compared to Wait trials from the same block (Wait-C and Wait-P, respectively). We omitted comparisons to Wait trials in the opposite block type (gray cells), which have little interpretability and were not addressed in our predictions. Asterisks mark statistically significant comparisons. D: AIC and BIC values for models of increasing complexity (AllMain refers to the a priori model which included main effects of cost condition and reward offer amount.)"}
# earnFitplot_wth, hrPlot_wth, mixLogis_compare_wth$plot
#ggsave("./Images/aic_wth.png", mixLogis_compare_wth$plot, device = "png", height = 5, width = 5)
include_graphics("./Images/FIG6_wthlogistic.png")
```

@@

```{r New Survival and prepost plots (Fig. 7), eval = figsEnd, out.width = "95%", fig.align = "center", fig.cap = "Choice consistency and stability in Experiment 2. A: Survival curves showing trial rejection times for each cost condition. B: Distribution of trial rejection times within the offer window. C: Acceptance rate in the last two blocks as a function of acceptance rate in the first two blocks. Dashed lines indicate the mean for each cost. D: Acceptance rates in the three successive occurrences of each block type."}
# plots generated in 16.3.1. and 16.3.3.
# historyPlot_wth <- (suppressWarnings(survPlot_wth) | suppressWarnings(RTplot_wth)) / (prepost_wth$scatterplot | prepost_wth$propBlocks) +
#     plot_annotation(tag_levels = 'A') &
#     theme(plot.tag = element_text(size = 20, face = "bold"))
#ggsave("./Images/FIG7_survivalwth.png", historyPlot_wth, device = "png", height = 8, width = 10)
include_graphics("./Images/FIG7_survivalwth.png")
```

@@

```{r New BIC and parameter comparison (Fig. 9), eval = figsEnd, fig.align = "center", out.width = "95%", fig.cap = "Computational modeling results. A: iBIC values for four candidate models of choice behavior for both experiments. Lower values indicate better fit. B: Estimates of the bias parameter c-demand for participants in the Physical, Cognitive, and Wait cost conditions of Experiment 1. Negative values suggest the Cognitive group acted as if cognitive demands attenuated the perceived opportunity cost of the delay. C: Estimates of other model parameters indicated consistent foraging behavior across conditions and experiments. D: Synthetic behavior for Experiment 2 generated by the winning model. Error bars correspond to the empirical mean acceptance rate ± SEM per cost type, whereas lines and shaded areas show the model-generated mean acceptance rate and its SEM, respectively. The model's behavior recapitulated the initial apparent preference for cognitive effort and the eventual convergence of acceptance rates by the end of the session."}
# modelPlot <- (bicPlot + costPlot) / allFitsPlot
# ggsave("./Images/FIG9_modelresult_part.png", modelPlot, device = "png", height = 8, width = 11)
include_graphics("./Images/FIG9_modelresults.png")
```
